Convex Surrogates via Polyhedral Losses

### 2019-02-01 cvx-fin.tex
Proofreading appendix

  We have now shown that $\gamma$ is $n$-embeddable.
  To drop a dimension, first note that $C$ is linear in the $\ones$ direction: for all $u'\in\reals^\Y$, the directional derivative $C'(u';\ones) = \sup_{p\in\partial C(u')} \inprod{p}{\ones} = 1$ as $\partial C(u')\subseteq\simplex$ for all $u'$; thus
  $C(u+\alpha\ones) = C(u) + \int_0^\alpha C'(u+\beta\ones;\ones) d\beta = C(u) + \alpha$.
  This linearity translates to invariance of $L$ in the direction of $\ones$:
  \begin{align*}
    L(u + \alpha\ones) = C(u+\alpha\ones)\ones - (u + \alpha\ones) = C(u)\ones + \alpha\ones - u - \alpha\ones = L(u)~.
  \end{align*}
  We now have for any $u\in\D$,
  \begin{align*}
    s^{-1}(u) \in \gamma(p) \iff u \in \prop{L}(p) \iff u - u_n\ones \in \prop{L}(p)\cap\R_2~.
  \end{align*}
  Letting $L':\reals^{n-1}\to\reals^\Y$ given by $L'(u') = L((u',0))$, and evoking Lemma~\ref{lem:loss-restrict}, we conclude $\prop{L'}(p) = \prop{L}\cap\R_2$ and thus $L'$ embeds $\gamma$ as well, with the embedding $\varphi(r) = (s(r)_1 - s(r)_n, \ldots, s(r)_{n-1} - s(r)_n)$.
\end{proof}

@@

  For the converse, let $\gamma:\simplex\toto\R$ be finite and orderable.
  From~\cite[Theorem 4]{lambert2018elicitation}, we have positively-oriented normals $v_i\in\reals^\Y$ for all $i \in \{1,\ldots,k-1\}$ such that $\gamma_{r_i} \cap \gamma_{r_{i+1}} = \{p\in\simplex : \inprod{v_i}{p} = 0\}$, and moreover, for all $i \in \{2,\ldots,k-1\}$, we have $\gamma_{r_i} = \{p\in\simplex : \inprod{v_{i-1}}{p} \leq 0 \leq \inprod{v_i}{p}\}$, while $\gamma_{r_1} = \{p\in\simplex : 0 \leq \inprod{v_1}{p} \}$ and $\gamma_{r_k} = \{p\in\simplex : \inprod{v_{k-1}}{p} \leq 0\}$.
  From the positive orientation of the $v_i$, we have for all $p\in\simplex$ that $\sgn(\inprod{v_i}{p})$ is monotone in $i$.
  In particular, it must be that for all $y$, $\sgn((v_i)_y)$ is monotone in $i$, taking the distribution with all weight on outcome $y$.
  % (Similarly, if $\gamma(\ones_y) = \{r_j,r_{j+1}\}$, then $(v_i)_y = v_i \cdot \ones_y < 0$ for $i < j$, $(v_j)_y = 0$, and $(v_i)_y > 0$ for $i > j$.)
  \raft{This observation could use a better proof.}

  For all $i\in\{2,\ldots,k-1\}$, we wish to find $\alpha_i \geq 0$ such that $v_{i-1} \leq \alpha_i v_i$ (component-wise).
  To this end, fix $i$ and let $v = v_{i-1}, w = v_{i}, \alpha=\alpha_i$.
  By the above, there is no $y\in\Y$ with $v_y > 0 > w_y$, so the condition $v \leq \alpha w$ is equivalent to the following:
  \begin{enumerate}
  \item[(i)] For all $y$ with $v_y,w_y < 0$, $\alpha \leq v_y/w_y$.
  \item[(ii)] For all $y$ with $v_y,w_y > 0$, $\alpha \geq v_y/w_y$.
  \end{enumerate}
  Defining $\Y^- = \{y\in\Y: v_i,w_y < 0\}, \Y^+ = \{y\in\Y: v_i,w_y > 0\}$, these conditions are in turn equivalent to $\min_{y\in\Y^-} v_y/w_y \geq \max_{y\in\Y^+} v_y/w_y$, with the usual conventions $\min \emptyset = \infty,\; \max \emptyset = -\infty$.
  Suppose this inequality were not satisfied.
  Then we would have $y \in \Y^-,y'\in\Y^+$ such that $0 < v_y/w_y < w_{y'}/v_y$, which would in turn imply $|v_y|/v_{y'} < |w_y| / w_{y'}$.
  Letting $c = \tfrac 1 2 \left(|w_y| / w_{y'} + |v_y|/v_{y'}\right)$ and taking $p$ to be the distribution with weight $1/(1+c)$ on $y$ and $c/(1+c)$ on $y'$, we see that
  \begin{align*}
    \inprod{v}{p} &= \frac 1 {1+c} \left(v_y + \tfrac 1 2 (|w_y| / w_{y'} + |v_y|/v_{y'})v_{y'}\right) > \frac 1 {1+c} \left(v_y + (|v_y|/v_{y'})v_{y'}\right) = 0
    \\
    \inprod{w}{p} &= \frac 1 {1+c} \left(w_y + \tfrac 1 2 (|w_y| / w_{y'} + |v_y|/v_{y'})v_{y'}\right) < \frac 1 {1+c} \left(w_y + (|w_y|/w_{y'})w_{y'}\right) = 0~,
  \end{align*}
  thus violating the observation that $\sgn(\inprod{v_i}{p})$ is monotone in $i$ (recall that $v = v_{i-1}, w = v_{i}$).

>>>>

  We have now shown that $\gamma$ is $n$-embeddable.
  To drop a dimension, first note that $C$ is linear in the $\ones$ direction: for all $u'\in\reals^\Y$, the directional derivative $C'(u';\ones) = \sup_{p\in\partial C(u')} \inprod{p}{\ones} = 1$ as $\partial C(u')\subseteq\simplex$ for all $u'$; thus
  $C(u+\alpha\ones) = C(u) + \int_0^\alpha C'(u+\beta\ones;\ones) d\beta = C(u) + \alpha$.
  This linearity translates to invariance of $L$ in the direction of $\ones$:
  \begin{align*}
    L(u + \alpha\ones) = C(u+\alpha\ones)\ones - (u + \alpha\ones) = C(u)\ones + \alpha\ones - u - \alpha\ones = L(u)~.
  \end{align*}
  We now have for any $u\in\D$,
  \begin{align*}
    s^{-1}(u) \in \gamma(p) \iff u \in \prop{L}(p) \iff u - u_n\ones \in \prop{L}(p)\cap\R_2~.
  \end{align*}
  Letting $L':\reals^{n-1}\to\reals^\Y$ given by $L'(u') = L((u',0))$, and evoking Lemma~\ref{lem:loss-restrict}, we conclude $\prop{L'}(p) = \prop{L}\cap\R_2$ and thus $L'$ embeds $\gamma$ as well, with the embedding $\varphi(r) = (s(r)_1 - s(r)_n, \ldots, s(r)_{n-1} - s(r)_n)$.
  \jessiet{The notation $C'$ being a derivative kind of bothers me since $L'$ is used for another generic function... although I may just be misreading it or not familiar with standard notations.}
\end{proof}

@@

  For the converse, let $\gamma:\simplex\toto\R$ be finite and orderable.
  From~\cite[Theorem 4]{lambert2018elicitation}, we have positively-oriented normals $v_i\in\reals^\Y$ for all $i \in \{1,\ldots,k-1\}$ such that $\gamma_{r_i} \cap \gamma_{r_{i+1}} = \{p\in\simplex : \inprod{v_i}{p} = 0\}$, and moreover, for all $i \in \{2,\ldots,k-1\}$, we have $\gamma_{r_i} = \{p\in\simplex : \inprod{v_{i-1}}{p} \leq 0 \leq \inprod{v_i}{p}\}$, while $\gamma_{r_1} = \{p\in\simplex : 0 \leq \inprod{v_1}{p} \}$ and $\gamma_{r_k} = \{p\in\simplex : \inprod{v_{k-1}}{p} \leq 0\}$.
  \jessiet{Should the sign on $\gamma_{r_k}$ be changed?}
  From the positive orientation of the $v_i$, we have for all $p\in\simplex$ that $\sgn(\inprod{v_i}{p})$ is monotone in $i$.
  In particular, it must be that for all $y$, $\sgn((v_i)_y)$ is monotone in $i$, taking the distribution with all weight on outcome $y$.
  % (Similarly, if $\gamma(\ones_y) = \{r_j,r_{j+1}\}$, then $(v_i)_y = v_i \cdot \ones_y < 0$ for $i < j$, $(v_j)_y = 0$, and $(v_i)_y > 0$ for $i > j$.)
  \raft{This observation could use a better proof.}

  For all $i\in\{2,\ldots,k-1\}$, we wish to find $\alpha_i \geq 0$ such that $v_{i-1} \leq \alpha_i v_i$ (component-wise).
  \jessiet{How should $w$ be interpreted?}
  To this end, fix $i$ and let $v = v_{i-1}, w = v_{i}, \alpha=\alpha_i$.
  By the above, there is no $y\in\Y$ with $v_y > 0 > w_y$, so the condition $v \leq \alpha w$ is equivalent to the following:
  \begin{enumerate}
  \item[(i)] For all $y$ with $v_y,w_y < 0$, $\alpha \leq v_y/w_y$.
  \item[(ii)] For all $y$ with $v_y,w_y > 0$, $\alpha \geq v_y/w_y$.
  \end{enumerate}
  Defining $\Y^- = \{y\in\Y: v_i,w_y < 0\}, \Y^+ = \{y\in\Y: v_i,w_y > 0\}$, these conditions are in turn equivalent to $\min_{y\in\Y^-} v_y/w_y \geq \max_{y\in\Y^+} v_y/w_y$, with the usual conventions $\min \emptyset = \infty,\; \max \emptyset = -\infty$.
  Suppose this inequality were not satisfied.
  Then we would have $y \in \Y^-,y'\in\Y^+$ such that $0 < v_y/w_y < w_{y'}/v_y$, which would in turn imply $|v_y|/v_{y'} < |w_y| / w_{y'}$.
  Letting $c = \tfrac 1 2 \left(|w_y| / w_{y'} + |v_y|/v_{y'}\right)$ and taking $p$ to be the distribution with weight $1/(1+c)$ on $y$ and $c/(1+c)$ on $y'$, we see that
  \begin{align*}
    \inprod{v}{p} &= \frac 1 {1+c} \left(v_y + \tfrac 1 2 (|w_y| / w_{y'} + |v_y|/v_{y'})v_{y'}\right) > \frac 1 {1+c} \left(v_y + (|v_y|/v_{y'})v_{y'}\right) = 0
    \\
    \inprod{w}{p} &= \frac 1 {1+c} \left(w_y + \tfrac 1 2 (|w_y| / w_{y'} + |v_y|/v_{y'})v_{y'}\right) < \frac 1 {1+c} \left(w_y + (|w_y|/w_{y'})w_{y'}\right) = 0~,
  \end{align*}
  thus violating the observation that $\sgn(\inprod{v_i}{p})$ is monotone in $i$ (recall that $v = v_{i-1}, w = v_{i}$).
